---
title: "Data Sourcing"
author: "Sam Pastoriza, Haley Roberts, Hyukzoo Shin, Harrison Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    self_contained: yes
    toc: yes
    number_sections: yes
  pdf_document:
    toc: yes
    keep_tex: yes
  word_document:
    toc: yes
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{tabulary}
  - \usepackage{multirow}
  - \usepackage{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

# Support Packages

```{r support, echo=FALSE,warning=FALSE,message=FALSE}
set.seed(114)
require(MASS)
require(glmnet)
library(kableExtra)
library(stargazer)
library(flextable)
library(dplyr)
library(leaps)
library(caret)
library(ggplot2)
library(ggExtra)
library(tidyr)
library(ISLR2)
library(ISLR)
library(readr)
library(reticulate)
library(tidyverse)
if (!require("pacman")) install.packages("pacman")
pacman::p_load("weights", "interactions", "cjoint", "plm", "interactions", "jtools", "stats", "miceadds", "broom", "RColorBrewer", "ggstatsplot", "ggpubr", "stargazer", "sandwich", "hrbrthemes", "rms", "interplot", "coefplot", "gmodels", "car", "lattice", "foreign", "ggplot2", "MASS", "Hmisc", "reshape2", "oddsratio", "tidyr", "psych", "dplyr", "tidyverse", "cjoint", "naniar", "scales")

knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE, echo = TRUE)

def.chunk.hook <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size, "\n\n", x, "\n\n \\normalsize"), x)
})

theme_earth <- function() {
  theme() %+replace%
    
    theme(
      panel.background = element_rect(fill = "white",
                                      colour = "white"),
      plot.title = element_text(
        size = 18,
        face = "bold",
        hjust = 0.5,
        #vjust = 2
      ),
      plot.subtitle = element_text(
        size = 12,
        hjust = 0.5,
      ),
      plot.caption = element_text(
        size = 10,
        hjust = 0.5
      ),
      axis.title = element_text(
        face = "bold",
        size = 14,
        hjust = 0.5
      ),
      axis.text = element_text(
        size = 12
      ),
      strip.text.x = element_text(
        size = 13
      )
    )
}
```

# Data Sourcing

## Load the training values
```{r}
training.values.df <- read.csv('./data/train_values.csv')
head(training.values.df)
tail(training.values.df)
str(training.values.df)
```

## Load the training labels

```{r}
training.labels.df <- read.csv('./data/train_labels.csv')
head(training.labels.df)
tail(training.labels.df)
str(training.labels.df)
```

# Merge the training data with the labels

```{r}
training.df <- training.values.df %>%
  left_join(training.labels.df, by = "building_id") %>%
  mutate(
    damage_grade = factor(damage_grade, levels = c(1, 2, 3), labels = c("Low", "Medium", "High")),
    land_surface_condition = as.factor(land_surface_condition),
    foundation_type = as.factor(foundation_type),
    roof_type = as.factor(roof_type),
    ground_floor_type = as.factor(ground_floor_type),
    other_floor_type = as.factor(other_floor_type),
    position = as.factor(position),
    plan_configuration = as.factor(plan_configuration),
    legal_ownership_status = as.factor(legal_ownership_status)
  )
training.df <- training.df[c(1, 40, 2:39)]

str(training.df)
head(training.df)

write.csv(training.df, './cleaned_data/cleaned_training_data.csv', row.names = FALSE)

training.numeric.df <- training.df %>% 
  mutate(land_surface_condition_n = ifelse(land_surface_condition == 'n', 1, 0),
         land_surface_condition_o = ifelse(land_surface_condition == 'o', 1, 0),
         land_surface_condition_t = ifelse(land_surface_condition == 't', 1, 0),
         foundation_type_h = ifelse(foundation_type == 'h', 1, 0),
         foundation_type_i = ifelse(foundation_type == 'i', 1, 0),
         foundation_type_r = ifelse(foundation_type == 'r', 1, 0),
         foundation_type_u = ifelse(foundation_type == 'u', 1, 0),
         foundation_type_w = ifelse(foundation_type == 'w', 1, 0),
         roof_type_n = ifelse(roof_type == 'n', 1, 0),
         roof_type_q = ifelse(roof_type == 'q', 1, 0),
         roof_type_x = ifelse(roof_type == 'x', 1, 0),
         ground_floor_type_f = ifelse(ground_floor_type == 'f', 1, 0),
         ground_floor_type_m = ifelse(ground_floor_type == 'm', 1, 0),
         ground_floor_type_v = ifelse(ground_floor_type == 'v', 1, 0),
         ground_floor_type_x = ifelse(ground_floor_type == 'x', 1, 0),
         ground_floor_type_z = ifelse(ground_floor_type == 'z', 1, 0),
         other_floor_type_j = ifelse(other_floor_type == 'j', 1, 0),
         other_floor_type_q = ifelse(other_floor_type == 'q', 1, 0),
         other_floor_type_s = ifelse(other_floor_type == 's', 1, 0),
         other_floor_type_x = ifelse(other_floor_type == 'x', 1, 0),
         position_j = ifelse(position == 'j', 1, 0),
         position_o = ifelse(position == 'o', 1, 0),
         position_s = ifelse(position == 's', 1, 0),
         position_t = ifelse(position == 't', 1, 0),
         plan_configuration_a = ifelse(plan_configuration == 'a', 1, 0),
         plan_configuration_c = ifelse(plan_configuration == 'c', 1, 0),
         plan_configuration_d = ifelse(plan_configuration == 'd', 1, 0),
         plan_configuration_f = ifelse(plan_configuration == 'f', 1, 0),
         plan_configuration_m = ifelse(plan_configuration == 'm', 1, 0),
         plan_configuration_n = ifelse(plan_configuration == 'n', 1, 0),
         plan_configuration_o = ifelse(plan_configuration == 'o', 1, 0),
         plan_configuration_q = ifelse(plan_configuration == 'q', 1, 0),
         plan_configuration_s = ifelse(plan_configuration == 's', 1, 0),
         plan_configuration_u = ifelse(plan_configuration == 'u', 1, 0),
         legal_ownership_status_a = ifelse(legal_ownership_status == 'a', 1, 0),
         legal_ownership_status_r = ifelse(legal_ownership_status == 'r', 1, 0),
         legal_ownership_status_v = ifelse(legal_ownership_status == 'v', 1, 0),
         legal_ownership_status_w = ifelse(legal_ownership_status == 'w', 1, 0)) %>%
  select(-land_surface_condition, -foundation_type, -roof_type, -ground_floor_type, -other_floor_type, -position, -plan_configuration, -legal_ownership_status)
  
str(training.numeric.df)
write.csv(training.numeric.df, './cleaned_data/cleaned_numeric_training_data.csv', row.names = FALSE)

training.df <- training.df %>%
  mutate(
    has_secondary_use = as.factor(has_secondary_use),
    has_secondary_use_agriculture = as.factor(has_secondary_use_agriculture),
    has_secondary_use_hotel = as.factor(has_secondary_use_hotel),
    has_secondary_use_rental = as.factor(has_secondary_use_rental),
    has_secondary_use_institution = as.factor(has_secondary_use_institution),
    has_secondary_use_school = as.factor(has_secondary_use_school),
    has_secondary_use_industry = as.factor(has_secondary_use_industry),
    has_secondary_use_health_post = as.factor(has_secondary_use_health_post),
    has_secondary_use_gov_office = as.factor(has_secondary_use_gov_office),
    has_secondary_use_use_police = as.factor(has_secondary_use_use_police),
    has_secondary_use_other = as.factor(has_secondary_use_other),
    has_superstructure_adobe_mud = as.factor(has_superstructure_adobe_mud),
    has_superstructure_mud_mortar_stone = as.factor(has_superstructure_mud_mortar_stone),
    has_superstructure_stone_flag = as.factor(has_superstructure_stone_flag),
    has_superstructure_cement_mortar_stone = as.factor(has_superstructure_cement_mortar_stone),
    has_superstructure_mud_mortar_brick = as.factor(has_superstructure_mud_mortar_brick),
    has_superstructure_cement_mortar_brick = as.factor(has_superstructure_cement_mortar_brick),
    has_superstructure_timber = as.factor(has_superstructure_timber),
    has_superstructure_bamboo = as.factor(has_superstructure_bamboo),
    has_superstructure_rc_non_engineered = as.factor(has_superstructure_rc_non_engineered),
    has_superstructure_rc_engineered = as.factor(has_superstructure_rc_engineered),
    has_superstructure_other = as.factor(has_superstructure_other))
```

# Transform the test data in the same fashion

```{r}
test.values.df <- read.csv('./data/test_values.csv')
head(test.values.df)
tail(test.values.df)
str(test.values.df)

write.csv(test.values.df, './cleaned_data/cleaned_testing_data.csv', row.names = FALSE)

test.numeric.df <- test.values.df %>% 
  mutate(land_surface_condition_n = ifelse(land_surface_condition == 'n', 1, 0),
         land_surface_condition_o = ifelse(land_surface_condition == 'o', 1, 0),
         land_surface_condition_t = ifelse(land_surface_condition == 't', 1, 0),
         foundation_type_h = ifelse(foundation_type == 'h', 1, 0),
         foundation_type_i = ifelse(foundation_type == 'i', 1, 0),
         foundation_type_r = ifelse(foundation_type == 'r', 1, 0),
         foundation_type_u = ifelse(foundation_type == 'u', 1, 0),
         foundation_type_w = ifelse(foundation_type == 'w', 1, 0),
         roof_type_n = ifelse(roof_type == 'n', 1, 0),
         roof_type_q = ifelse(roof_type == 'q', 1, 0),
         roof_type_x = ifelse(roof_type == 'x', 1, 0),
         ground_floor_type_f = ifelse(ground_floor_type == 'f', 1, 0),
         ground_floor_type_m = ifelse(ground_floor_type == 'm', 1, 0),
         ground_floor_type_v = ifelse(ground_floor_type == 'v', 1, 0),
         ground_floor_type_x = ifelse(ground_floor_type == 'x', 1, 0),
         ground_floor_type_z = ifelse(ground_floor_type == 'z', 1, 0),
         other_floor_type_j = ifelse(other_floor_type == 'j', 1, 0),
         other_floor_type_q = ifelse(other_floor_type == 'q', 1, 0),
         other_floor_type_s = ifelse(other_floor_type == 's', 1, 0),
         other_floor_type_x = ifelse(other_floor_type == 'x', 1, 0),
         position_j = ifelse(position == 'j', 1, 0),
         position_o = ifelse(position == 'o', 1, 0),
         position_s = ifelse(position == 's', 1, 0),
         position_t = ifelse(position == 't', 1, 0),
         plan_configuration_a = ifelse(plan_configuration == 'a', 1, 0),
         plan_configuration_c = ifelse(plan_configuration == 'c', 1, 0),
         plan_configuration_d = ifelse(plan_configuration == 'd', 1, 0),
         plan_configuration_f = ifelse(plan_configuration == 'f', 1, 0),
         plan_configuration_m = ifelse(plan_configuration == 'm', 1, 0),
         plan_configuration_n = ifelse(plan_configuration == 'n', 1, 0),
         plan_configuration_o = ifelse(plan_configuration == 'o', 1, 0),
         plan_configuration_q = ifelse(plan_configuration == 'q', 1, 0),
         plan_configuration_s = ifelse(plan_configuration == 's', 1, 0),
         plan_configuration_u = ifelse(plan_configuration == 'u', 1, 0),
         legal_ownership_status_a = ifelse(legal_ownership_status == 'a', 1, 0),
         legal_ownership_status_r = ifelse(legal_ownership_status == 'r', 1, 0),
         legal_ownership_status_v = ifelse(legal_ownership_status == 'v', 1, 0),
         legal_ownership_status_w = ifelse(legal_ownership_status == 'w', 1, 0)) %>%
  select(-land_surface_condition, -foundation_type, -roof_type, -ground_floor_type, -other_floor_type, -position, -plan_configuration, -legal_ownership_status)

str(test.numeric.df)

write.csv(test.numeric.df, './cleaned_data/cleaned_numeric_testing_data.csv', row.names = FALSE)
```

# Initial Data Exploration

Using a simple missing values chart, we can see how many values are actually missing from the initial dataset.

```{r}
gg_miss_var(training.df %>% mutate_all(na_if, ""), show_pct = TRUE) +
  ggtitle('Missing Values Plot')
ggsave("./visualizations/missing_variables_plot.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/missing_variables_plot.png", width = 10, height = 8, units = "in")
```

Given the missing values chart above, there are no missing values across all of the variables.

# Learn more about categorical variable distributions

Plot distributions of categorical variables

```{r fig.height=8, fig.width=10}
training.df %>%
  select(land_surface_condition, 
         foundation_type, 
         roof_type, 
         ground_floor_type, 
         other_floor_type, 
         position, 
         plan_configuration, 
         legal_ownership_status) %>%
  pivot_longer(names_to = 'Category', values_to = 'Value', cols = everything()) %>% 
  ggplot(aes(x = Value)) + 
  geom_bar(stat = 'count') + 
  facet_wrap(~Category, scales = 'free') +
  labs(x = 'Variable', y = 'Count', title = 'Distribution of Categorial Variables') + 
  theme_earth()
ggsave("./visualizations/categorical_variables_distribution.svg", width = 10, height = 6, units = "in")
ggsave("./visualizations/categorical_variables_distribution.png", width = 10, height = 6, units = "in")
```

### Numerical Distribution

```{r fig.height=8, fig.width=10}
training.df %>%
  select(where(is.numeric)) %>%
  select(-building_id) %>%
  pivot_longer(names_to = "Variable", values_to = "value", cols = everything()) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~Variable, scales = "free") +
  labs(x = 'Variable', y = 'Count', title = 'Distribution of Numerical Variables') + 
  theme_earth()
ggsave("./visualizations/numerical_variables_distribution.svg", width = 10, height = 6, units = "in")
ggsave("./visualizations/numerical_variables_distribution.png", width = 10, height = 6, units = "in")
```

### Superstructure information

```{r fig.height=8, fig.width=10}
training.df %>%
  select(has_superstructure_adobe_mud, 
         has_superstructure_mud_mortar_stone, 
         has_superstructure_stone_flag, 
         has_superstructure_cement_mortar_stone, 
         has_superstructure_mud_mortar_brick, 
         has_superstructure_cement_mortar_brick, 
         has_superstructure_timber, 
         has_superstructure_bamboo, 
         has_superstructure_rc_non_engineered, 
         has_superstructure_rc_engineered,
         has_superstructure_other) %>%
  pivot_longer(names_to = "Category", values_to = "Value", cols = everything()) %>%
  mutate(Category = str_replace(Category, 'has_superstructure_', '')) %>%
  ggplot(aes(x = Value, fill = Value)) +
  geom_bar(stat = "count") +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1) +
  facet_wrap(~Category, scales = "free") +
  labs(x = 'Value', y = 'Count', title = 'Distribution of Superstructure Variables') + 
  theme_earth()
ggsave("./visualizations/superstructure_variables_distribution.svg", width = 14, height = 6, units = "in")
ggsave("./visualizations/superstructure_variables_distribution.png", width = 14, height = 6, units = "in")
```

Visualizing superstructure variable vs damage grade
```{r fig.height=8, fig.width=10}
training.df %>%
  select(has_superstructure_cement_mortar_brick, 
         has_superstructure_adobe_mud, 
         has_superstructure_mud_mortar_stone, 
         has_superstructure_stone_flag, 
         has_superstructure_cement_mortar_stone, 
         has_superstructure_mud_mortar_brick, 
         has_superstructure_cement_mortar_brick, 
         has_superstructure_timber, 
         has_superstructure_bamboo, 
         has_superstructure_rc_non_engineered, 
         has_superstructure_rc_engineered,
         has_superstructure_other,
         damage_grade) %>%
  pivot_longer(names_to = "Category", values_to = "Value", cols = c(-damage_grade)) %>%
  mutate(Category = str_replace(Category, 'has_superstructure_', '')) %>%
  filter(Value == 1) %>%
  ggplot(aes(x = damage_grade, fill = damage_grade)) +
  geom_bar(stat = "count") +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1, nudge_y = -.5) + 
  facet_wrap(~Category, scales = "free") +
  labs(x = 'Damage Grade', y = 'Count', title = 'Superstructure Type vs Damage Sustained', subtitle = 'Only when a building has a superstructure type') + 
  guides(fill = guide_legend(title = "Damage Grade")) + 
  theme_earth()
ggsave("./visualizations/superstructure_damage_comparison.svg", width = 14, height = 6, units = "in")
ggsave("./visualizations/superstructure_damage_comparison.png", width = 14, height = 6, units = "in")
```


### Secondary information

```{r fig.height=8, fig.width=10}
training.df %>%
  select(
    has_secondary_use_agriculture,
    has_secondary_use_hotel,
    has_secondary_use_rental,
    has_secondary_use_institution,
    has_secondary_use_school,
    has_secondary_use_industry,
    has_secondary_use_health_post,
    has_secondary_use_gov_office,
    has_secondary_use_use_police,
    has_secondary_use_other
  ) %>%
  pivot_longer(names_to = "Category", values_to = "Value", cols = everything()) %>%
  mutate(Category = str_replace(Category, 'has_secondary_use_', '')) %>%
  ggplot(aes(x = Value, fill = Value)) +
  geom_bar(stat = "count") +
  geom_text(aes(label = ..count..), stat = "count") +
  facet_wrap(~Category, scales = "free") +
  labs(x = 'Value', y = 'Count', title = 'Distribution of Secondary Use Variables') + 
  theme_earth()
ggsave("./visualizations/secondary_variables_distribution.svg", width = 12, height = 6, units = "in")
ggsave("./visualizations/secondary_variables_distribution.png", width = 12, height = 6, units = "in")
```

Plot secondary usage vs damage grade
```{r}
training.df %>%
  select(has_secondary_use_agriculture,
         has_secondary_use_hotel,
         has_secondary_use_rental,
         has_secondary_use_institution,
         has_secondary_use_school,
         has_secondary_use_industry,
         has_secondary_use_health_post,
         has_secondary_use_gov_office,
         has_secondary_use_use_police,
         has_secondary_use_other,
         damage_grade) %>%
  pivot_longer(names_to = "Category", values_to = "Value", cols = c(-damage_grade)) %>%
  mutate(Category = str_replace(Category, 'has_secondary_use_', '')) %>%
  filter(Value == 1) %>%
  ggplot(aes(x = damage_grade, fill = damage_grade)) +
  geom_bar(stat = "count") +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1, nudge_y = -.5) + 
  facet_wrap(~Category, scales = "free") +
  labs(x = 'Damage Grade', y = 'Count', title = 'Secondary Use vs Damage Sustained', subtitle = 'Only when a building has a secondary use') + 
  guides(fill = guide_legend(title = "Damage Grade")) + 
  theme_earth()
ggsave("./visualizations/secondary_use_damage_comparison.svg", width = 14, height = 6, units = "in")
ggsave("./visualizations/secondary_use_damage_comparison.png", width = 14, height = 6, units = "in")
```

Plot damage grade label information

```{r}
training.df %>%
  ggplot(aes(x = damage_grade, fill = damage_grade)) +
  geom_bar() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, nudge_y = -.5) +
  guides(color = guide_legend('Damage Grade')) + 
  labs(x = 'Damage Grade', y = 'Count', title = 'Distribution of Damage Grades') + 
  guides(fill = guide_legend(title = "Damage Grade")) + 
  theme_earth()
ggsave("./visualizations/labels_distribution.svg", width = 10, height = 6, units = "in")
ggsave("./visualizations/labels_distribution.png", width = 10, height = 6, units = "in")
```

## Plot Correlation Matrices

```{r fig.height=8, fig.width=10}
# GGally::ggpairs(training.df[,2:6], 
#                 aes(color = training.df$damage_grade, alpha = 0.5), 
#                 title = 'Correlogram of Data') 
```


```{r fig.height=8, fig.width=10}
# GGally::ggpairs(training.df[,17:26], 
#                 aes(color = training.df$damage_grade, alpha = 0.5), 
#                 title = 'Correlogram of Superstructure Data') 
```

# Check the categorical variables compared to damage grade

## Land Surface Condition

```{r}
land.heatmap.df <- as.data.frame(table(training.df$land_surface_condition, training.df$damage_grade)) %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(land.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    geom_text(aes(label = paste0(NormalizedFreq, '%'))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Land Surface Condition", y = "Damage Type", title = "Land Surface Condition Heatmap -- Normalized") + 
    theme_earth()
ggsave("./visualizations/land_surface_condition_normalized_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/land_surface_condition_normalized_heatmap.png", width = 10, height = 8, units = "in")

ggplot(land.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    geom_text(aes(label = round(Freq, 1))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Land Surface Condition", y = "Damage Type", title = "Land Surface Condition Heatmap -- Not Normalized") + 
    theme_earth()
ggsave("./visualizations/land_surface_condition_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/land_surface_condition_heatmap.png", width = 10, height = 8, units = "in")
```

## Foundation Type

```{r}
foundation.heatmap.df <- as.data.frame(table(training.df$foundation_type, training.df$damage_grade)) %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(foundation.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    geom_text(aes(label = paste0(NormalizedFreq, '%'))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Foundation Type", y = "Damage Type", title = "Foundation Type Heatmap -- Normalized") + 
    theme_earth()
ggsave("./visualizations/foundation_type_normalized_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/foundation_type_normalized_heatmap.png", width = 10, height = 8, units = "in")

ggplot(foundation.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    geom_text(aes(label = round(Freq, 1))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Foundation Type", y = "Damage Type", title = "Foundation Type Heatmap -- Not Normalized") + 
    theme_earth()
ggsave("./visualizations/foundation_type_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/foundation_type_heatmap.png", width = 10, height = 8, units = "in")
```

## Roof Type

```{r}
roof.heatmap.df <- as.data.frame(table(training.df$roof_type, training.df$damage_grade)) %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(roof.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    geom_text(aes(label = paste0(NormalizedFreq, "%"))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Roof Type", y = "Damage Type", title = "Roof Type Heatmap -- Normalized") + 
    theme_earth()
ggsave("./visualizations/roof_type_normalized_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/roof_type_normalized_heatmap.png", width = 10, height = 8, units = "in")

ggplot(roof.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    geom_text(aes(label = round(Freq, 1))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Roof Type", y = "Damage Type", title = "Roof Type Heatmap -- Not Normalized") + 
    theme_earth()
ggsave("./visualizations/roof_type_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/roof_type_heatmap.png", width = 10, height = 8, units = "in")
```

## Ground Floor Type

```{r}
ground.heatmap.df <- as.data.frame(table(training.df$ground_floor_type, training.df$damage_grade)) %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(ground.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    geom_text(aes(label = paste0(NormalizedFreq, "%"))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Ground Type", y = "Damage Type", title = "Ground Floor Type Heatmap -- Normalized")
ggsave("./visualizations/ground_floor_type_normalized_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/ground_floor_type_normalized_heatmap.png", width = 10, height = 8, units = "in")

ggplot(ground.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    geom_text(aes(label = round(Freq, 1))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Ground Type", y = "Damage Type", title = "Ground Floor Type Heatmap -- Not Normalized")
ggsave("./visualizations/ground_floor_type_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/ground_floor_type_heatmap.png", width = 10, height = 8, units = "in")
```

## Other Floor Type

```{r}
other.heatmap.df <- as.data.frame(table(training.df$other_floor_type, training.df$damage_grade)) %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(other.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    geom_text(aes(label = paste0(NormalizedFreq, "%"))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Other Floor Type", y = "Damage Type", title = "Other Floor Type Heatmap -- Normalized") + 
    theme_earth()
ggsave("./visualizations/other_floor_type_normalized_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/other_floor_type_normalized_heatmap.png", width = 10, height = 8, units = "in")

ggplot(other.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    geom_text(aes(label = round(Freq, 1))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Other Floor Type", y = "Damage Type", title = "Other Floor Type Heatmap -- Not Normalized") + 
    theme_earth()
ggsave("./visualizations/other_floor_type_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/other_floor_type_heatmap.png", width = 10, height = 8, units = "in")
```

## Plan Configuration

```{r}
plan.heatmap.df <- as.data.frame(table(training.df$plan_configuration, training.df$damage_grade)) %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(plan.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    geom_text(aes(label = paste0(NormalizedFreq, "%"))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Plan Configuration", y = "Damage Type", title = "Plan Configuration Heatmap -- Normalized") + 
    theme_earth()
ggsave("./visualizations/plan_configuration_normalized_heatmap.svg", width = 12, height = 8, units = "in")
ggsave("./visualizations/plan_configuration_normalized_heatmap.png", width = 12, height = 8, units = "in")

ggplot(plan.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    geom_text(aes(label = round(Freq, 1))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Plan Configuration", y = "Damage Type", title = "Plan Configuration Heatmap -- Not Normalized") + 
    theme_earth()
ggsave("./visualizations/plan_configuration_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/plan_configuration_heatmap.png", width = 10, height = 8, units = "in")
```

## Legal Ownership Status

```{r}
legal.heatmap.df <- as.data.frame(table(training.df$legal_ownership_status, training.df$damage_grade)) %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(legal.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    geom_text(aes(label = paste0(NormalizedFreq, "%"))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Legal Ownership Status", y = "Damage Type", title = "Legal Ownership Status Heatmap -- Normalized") + 
    theme_earth()
ggsave("./visualizations/legal_ownership_status_normalized_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/legal_ownership_status_normalized_heatmap.png", width = 10, height = 8, units = "in")

ggplot(legal.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    geom_text(aes(label = round(Freq, 1))) +
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Legal Ownership Status", y = "Damage Type", title = "Legal Ownership Status Heatmap -- Not Normalized") + 
    theme_earth()
ggsave("./visualizations/legal_ownership_status_heatmap.svg", width = 10, height = 8, units = "in")
ggsave("./visualizations/legal_ownership_status_heatmap.png", width = 10, height = 8, units = "in")
```

# Outlier Checking

What is going on with the Age? 995 for age?
```{r}
table(training.df$age)

age.heatmap.df <- as.data.frame(table(training.df$age, training.df$damage_grade)) %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(age.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Age", y = "Damage Type", title = "Age Heatmap -- Normalized") + 
    theme_earth()
ggsave("./visualizations/age_normalized_heatmap.svg", width = 16, height = 6, units = "in")
ggsave("./visualizations/age_normalized_heatmap.png", width = 16, height = 6, units = "in")

ggplot(age.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    scale_fill_gradient(low = "white", high = "red")  + 
    labs(x = "Age", y = "Damage Type", title = "Age Heatmap -- Not Normalized") + 
    theme_earth()
ggsave("./visualizations/age_heatmap.svg", width = 16, height = 6, units = "in")
ggsave("./visualizations/age_heatmap.png", width = 16, height = 6, units = "in")
```

# Feature Engineering

### Feature Engineering (Secondary Variables)

Checking to see if we can combine secondary use data into one variable

```{r}
secondary.df <- training.df %>%
  mutate(secondary_use = ifelse(has_secondary_use_agriculture == 1, "agriculture", "")) %>%
  mutate(secondary_use = ifelse(has_secondary_use_hotel == 1, paste(secondary_use, "hotel", sep = " "), secondary_use)) %>%
  mutate(secondary_use = ifelse(has_secondary_use_rental == 1, paste(secondary_use, "rental", sep = " "), secondary_use)) %>%
  mutate(secondary_use = ifelse(has_secondary_use_institution == 1, paste(secondary_use, "institution", sep = " "), secondary_use)) %>%
  mutate(secondary_use = ifelse(has_secondary_use_school == 1, paste(secondary_use, "school", sep = " "), secondary_use)) %>%
  mutate(secondary_use = ifelse(has_secondary_use_industry == 1, paste(secondary_use, "industry", sep = " "), secondary_use)) %>%
  mutate(secondary_use = ifelse(has_secondary_use_health_post == 1, paste(secondary_use, "health_post", sep = " "), secondary_use)) %>%
  mutate(secondary_use = ifelse(has_secondary_use_gov_office == 1, paste(secondary_use, "gov_office", sep = " "), secondary_use)) %>%
  mutate(secondary_use = ifelse(has_secondary_use_use_police == 1, paste(secondary_use, "use_police", sep = " "), secondary_use)) %>%
  mutate(secondary_use = ifelse(has_secondary_use_other == 1, paste(secondary_use, "other", sep = " "), secondary_use)) %>%
  mutate(secondary_use = str_trim(secondary_use)) %>%
  mutate(secondary_use = ifelse(secondary_use == "", "none", secondary_use))

table(secondary.df$secondary_use)
table(secondary.df$secondary_use, secondary.df$damage_grade)

secondary.heatmap.df <- as.data.frame(table(secondary.df$secondary_use, secondary.df$damage_grade)) %>%
  filter(Var1 != 'none') %>%
  group_by(Var1) %>%
  mutate(TotalInVar = sum(Freq)) %>%
  ungroup() %>%
  mutate(NormalizedFreq = round((Freq / TotalInVar) * 100, digits = 2)) %>%
  mutate(NormalizedFreq = as.numeric(as.character(NormalizedFreq)))

ggplot(secondary.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = NormalizedFreq)) + 
    geom_text(aes(label = round(NormalizedFreq, 1))) +
    scale_fill_gradient(low = "white", 
                        high = "red")  + 
    labs(x = "Secondary Use", y = "Damage Type", title = "Secondary Use Heatmap -- Normalized") + 
    theme_earth()
ggsave("./visualizations/secondary_normalized_heatmap.svg", width = 16, height = 6, units = "in")
ggsave("./visualizations/secondary_normalized_heatmap.png", width = 16, height = 6, units = "in")

ggplot(secondary.heatmap.df, aes(Var1, Var2)) +
    geom_tile(aes(fill = Freq)) + 
    geom_text(aes(label = round(Freq, 1))) +
    scale_fill_gradient(low = "white", 
                        high = "red")  + 
    labs(x = "Secondary Use", y = "Damage Type", title = "Secondary Use Heatmap -- Not Normalized") + 
    theme_earth()
ggsave("./visualizations/secondary_heatmap.svg", width = 16, height = 6, units = "in")
ggsave("./visualizations/secondary_heatmap.png", width = 16, height = 6, units = "in")
```

This seems promising if we want to reduce the number of features.

### Feature Engineering (Superstructure Variables)

Checking to see if we can/want to combine superstructure data into one variable

```{r}
superstructure.df <- training.df %>%
    mutate(superstructure_material = ifelse(has_superstructure_adobe_mud == 1, "adobe_mud", "")) %>%
    mutate(superstructure_material = ifelse(has_superstructure_mud_mortar_stone == 1, paste(superstructure_material, "mud_mortar_stone", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_stone_flag == 1, paste(superstructure_material, "stone_flag", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_cement_mortar_stone == 1, paste(superstructure_material, "cement_mortar_stone", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_mud_mortar_brick == 1, paste(superstructure_material, "mud_mortar_brick", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_cement_mortar_brick == 1, paste(superstructure_material, "cement_mortar_brick", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_timber == 1, paste(superstructure_material, "timber", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_bamboo == 1, paste(superstructure_material, "bamboo", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_rc_non_engineered == 1, paste(superstructure_material, "rc_non_engineered", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_rc_engineered == 1, paste(superstructure_material, "rc_engineered", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = ifelse(has_superstructure_other == 1, paste(superstructure_material, "other", sep = " "), superstructure_material)) %>%
    mutate(superstructure_material = str_trim(superstructure_material)) %>%
    mutate(superstructure_material = ifelse(superstructure_material == "", "unknown", superstructure_material))

table(superstructure.df$superstructure_material)
```

This seems less promising as the variables are not independent of each other
